00-Data-Processing-for-sobiEquity

### 2021-03-17 Bike-share-spatial-equity/Bike-share-spatial-equity.Rmd
Added downtown areas to data package, and the SoBi hubs have a new identifier for status (some hubs have been deactivated). The code in the manuscript now filters non-active stations in the calculation of accessibility.

```{r}
data("hamilton_cma")
data("hamilton_da_2016")
data("population_50x50")
data("sobi_service")
data("sobi_hubs")
data("ttm_walk")
```

```{r coordinates-of-equity-stations, include=FALSE}
# define the coordinates of the equity stations
# equity_stations <- data.frame(longitude = c(-79.84364993846138, 
#                                             -79.8362839215856,
#                                             -79.83144521746212, 
#                                             -79.82217214949738, 
#                                             -79.8160211746482, 
#                                             -79.8192304372373, 
#                                             -79.82500255113169, 
#                                             -79.8268599805827, 
#                                             -79.83888834759689, 
#                                             -79.82740849222654, 
#                                             -79.82342541216097, 
#                                             -79.84240606419081), 
#                               latitude = c(43.265340158000654, 
#                                            43.25632467284136, 
#                                            43.25528943845642, 
#                                            43.25291050074978, 
#                                            43.25147804403345, 
#                                            43.24444659758296, 
#                                            43.2421627336722, 
#                                            43.24634382671116,
#                                            43.249432741404554,
#                                            43.251126444296524, 
#                                            43.24892849896579, 
#                                            43.25831920064561))
```

```{r, include=FALSE}
#equity_stations <- st_as_sf(equity_stations, coords = c("longitude", "latitude"), 
    #crs = st_crs(sobi_service), agr = "constant")
```

```{r load-data, include=FALSE}
#load("accessibility.RData") #full accessibility table for SoBi system
#load("accessibility_map_original.RData") #map of accessibility in original SoBi system
#load("accessibility_map_equity.RData") #map of accessibility in current system with ERI
# load("hamilton_cma.RData")
# load("hamilton_da_2016.RData")
# load("hamilton_da_sobi.RData")
# load("mhi.RData") #median household income by DA
# load("population_sobi.RData") #interpolated population in SoBi area
#load("population_sobi_original.RData") #population in original system
# load("sobi_hubs.RData") #location of SoBi hubs, minus Van Wagner's beach
# load("sobi_hubs_original.RData") #SoBi hubs, minus ERI hubs
# load("sobi_service.RData") #core service area for SoBi
# load("ttm_walk.RData") #travel time matrix for interpolated population in SoBi area
# load("ttm_walk_original.RData") #travel time matrix for system without ERI hubs
```

@@

```{r filter-eri-hubs, include=FALSE}
# eri_hubs <- sobi_hubs %>% 
#              filter(OBJECTID>=117 & OBJECTID<=128)
```

```{r, include=FALSE}
sobi_downtown <- st_read("Community_Downtown.shp")
```

```{r sobi-stations-in-hamilton, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "The spatial distribution of bike share stations in Hamilton, Ontario. The service area of Hamilton Bike Share is outlined in blue and the city's downtown core is outlined in light blue. Dissemination areas within the Hamilton CMA that touch the bounding box of the core service area are outlined in grey."}
ggplot() + 
  geom_sf(data = hamilton_da_2016 %>% 
            dplyr::filter(sobi_service == "Yes"),
          color = "gray") +
  geom_sf(data = sobi_service, 
          fill = NA,
          color = "black") +
  geom_sf(data = sobi_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = "dark green",
          color = "dark green") +
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "blue") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

@@

```{r figure-4, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population cells in the SoBi Hamilton core service area (outlined in black) and within 30 minutes of walking to the core service area (outlined in blue). The downtown area is outlined in dark green. Geographic features have been extracted. Dissemination area polygons are outlined in grey."}
ggplot() + 
  geom_sf(data = population_50x50,
          aes(color = population)) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = hamilton_da_2016 %>%
            filter(sobi_service == "Yes"),
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "black") +
  geom_sf(data = sobi_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green")
```

```{r figure-5, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population cells in the SoBi Hamilton core service area and within 30 minutes of walking to the core service area. Green circles represent the location of the equity stations that were added in 2018. Blue circles represent the conventional stations. The service area of the PBSP is outlined in black, and dissemination areas from the 2016 Canadian Census are outlined in light gray."}
ggplot() + 
  geom_sf(data = population_50x50, 
          aes(color = population)) +
  geom_sf(data = hamilton_da_2016 %>%
            dplyr::filter(sobi_service == "Yes"), 
          fill = NA, 
          color = "lightgray") + 
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "blue") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green") +
  geom_sf(data = sobi_service, 
          fill = NA, 
          colour = "black") +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = sobi_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  facet_grid(rows = vars(hub_type))
```

@@

```{r echo=FALSE}
#data frames for current and original systems
names(ttm_walk) <- c('UID', 
                     'hub', 
                     'travel_time',
                     "hub_type",
                     'population',
                     'racks')
```

```{r, echo=FALSE}
#population_sobi_original <- population_sobi_original[-c(3, 5:6)]
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 3 minutes:
```{r}
results_3min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 3)
```

All stations:
```{r}
results_3min_all <- ttm_walk %>%
  b2sfca(threshold = 3)
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 5 minutes:
```{r}
results_5min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 5)
```

All stations:
```{r}
results_5min_all <- ttm_walk %>%
  b2sfca(threshold = 5)
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 10 minutes:
```{r}
results_10min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 10)
```

All stations:
```{r}
results_10min_all <- ttm_walk %>%
  b2sfca(threshold = 10)
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 15 minutes:
```{r}
results_15min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 15)
```

All stations:
```{r}
results_15min_all <- ttm_walk %>%
  b2sfca(threshold = 15)
```

>>>>

```{r invoke-data, include=FALSE}
data("community_downtown")
data("hamilton_cma")
data("hamilton_da_2016")
data("population_50x50")
data("sobi_service")
data("sobi_hubs")
data("ttm_walk")
```

@@

```{r sobi-stations-in-hamilton, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "The spatial distribution of bike share stations in Hamilton, Ontario. The service area of Hamilton Bike Share is outlined in blue and the city's downtown core is outlined in light blue. Dissemination areas within the Hamilton CMA that touch the bounding box of the core service area are outlined in grey."}
ggplot() + 
  geom_sf(data = hamilton_da_2016 %>% 
            dplyr::filter(sobi_service == "Yes"),
          color = "gray") +
  geom_sf(data = sobi_service, 
          fill = NA,
          color = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = "dark green",
          color = "dark green") +
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional" & 
                            hub_status == "Active"), # Remove deactivated hubs
          size = 1.5, 
          shape = 21, 
          fill = "blue") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

@@

```{r figure-4, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population cells in the SoBi Hamilton core service area (outlined in black) and within 30 minutes of walking to the core service area (outlined in blue). The downtown area is outlined in dark green. Geographic features have been extracted. Dissemination area polygons are outlined in grey."}
ggplot() + 
  geom_sf(data = population_50x50,
          aes(color = population)) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = hamilton_da_2016 %>%
            filter(sobi_service == "Yes"),
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green")
```

```{r figure-5, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population cells in the SoBi Hamilton core service area and within 30 minutes of walking to the core service area. Green circles represent the location of the equity stations that were added in 2018. Blue circles represent the conventional stations. The service area of the PBSP is outlined in black, and dissemination areas from the 2016 Canadian Census are outlined in light gray."}
ggplot() + 
  geom_sf(data = population_50x50, 
          aes(color = population)) +
  geom_sf(data = hamilton_da_2016 %>%
            dplyr::filter(sobi_service == "Yes"), 
          fill = NA, 
          color = "lightgray") + 
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional" &
                            hub_status == "Active"), 
          size = 1.5, 
          shape = 21, 
          fill = "blue") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green") +
  geom_sf(data = sobi_service, 
          fill = NA, 
          colour = "black") +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  facet_grid(rows = vars(hub_type))
```

@@

```{r echo=FALSE}
#data frames for current and original systems
names(ttm_walk) <- c("UID", 
                     "hub", 
                     "travel_time",
                     "hub_type",
                     "hub_status",
                     "population",
                     "racks")
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 3 minutes:
```{r}
results_3min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

All stations:
```{r}
results_3min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 5 minutes:
```{r}
results_5min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

All stations:
```{r}
results_5min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 10 minutes:
```{r}
results_10min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

All stations:
```{r}
results_10min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

@@

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 15 minutes:
```{r}
results_15min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

All stations:
```{r}
results_15min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

### 2021-03-17 sobiEquity/raw-data-processing/00-Data-Processing-for-sobiEquity.Rmd
Added downtown areas to data package, and the SoBi hubs have a new identifier for status (some hubs have been deactivated). The code in the manuscript now filters non-active stations in the calculation of accessibility.

@@

@@

Plot again:
```{r}
ggplot() + 
  geom_sf(data = interpolated_pop_clean,
          aes(color = population)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  geom_sf(data = sobi_hubs,
          aes(shape = hub_type, 
              size = RACKS_AMOU)) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  scale_size(range = c(1, 4))
```

@@

@@

Join hub type information to travel time table:
```{r}
ttm_walk <- ttm_walk %>%
  left_join(sobi_hubs %>%
              st_drop_geometry() %>%
              dplyr::select(OBJECTID,
                            hub_type),
            by = "OBJECTID")
```

@@

Save data files:
```{r}
save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
save(hamilton_da_2016, file = "output-data-files/hamilton_da_2016.RData", compress = "xz")
save(population_50x50, file = "output-data-files/population_50x50.RData", compress = "xz")
save(sobi_hubs, file = "output-data-files/sobi_hubs.RData", compress = "xz")
save(sobi_service, file = "output-data-files/sobi_service.RData", compress = "xz")
save(ttm_walk, file = "output-data-files/ttm_walk.RData", compress = "xz")
```

>>>>

INFO

### Community Downtown

INFO

@@

Read community downtown:
```{r}
community_downtown <- st_read("input-data-files/Community_Downtown.shp")
```

@@

Plot again:
```{r}
ggplot() + 
  geom_sf(data = interpolated_pop_clean,
          aes(color = population)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  geom_sf(data = sobi_hubs,
          aes(size = RACKS_AMOU)) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  scale_size(range = c(1, 4))
```

@@

Identify hubs that were deactivated:
```{r}
sobi_hubs <- sobi_hubs %>%
  dplyr::mutate(hub_status = case_when(OBJECTID == 114 ~ "Deactivated",
                                       OBJECTID == 115 ~ "Deactivated",
                                       OBJECTID == 116 ~ "Deactivated",
                                       TRUE ~ "Active"),
                hub_status = factor(hub_status))

@@

Join hub type information to travel time table:
```{r}
ttm_walk <- ttm_walk %>%
  left_join(sobi_hubs %>%
              st_drop_geometry() %>%
              dplyr::select(OBJECTID,
                            hub_type,
                            hub_status),
            by = "OBJECTID")
```

@@

Save data files:
```{r}
save(community_downtown, file = "output-data-files/community_downtown.RData", compress = "xz")
save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
save(hamilton_da_2016, file = "output-data-files/hamilton_da_2016.RData", compress = "xz")
save(population_50x50, file = "output-data-files/population_50x50.RData", compress = "xz")
save(sobi_hubs, file = "output-data-files/sobi_hubs.RData", compress = "xz")
save(sobi_service, file = "output-data-files/sobi_service.RData", compress = "xz")
save(ttm_walk, file = "output-data-files/ttm_walk.RData", compress = "xz")
```
